---
title: "Custom Templates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom Templates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup}
library(thoth)
```

## Introduction

Quarto templates help maintain consistent branding and styling across your organization's reports. `thoth` provides tools to create and manage custom Quarto templates with your organization's branding, colors, and styling preferences.

## Basic Template Creation

Create a new template with your organization's branding:

```{r eval=FALSE}
library(thoth)

create_quarto_template(
  template_name = "company_template",
  primary_color = "#0054AD",
  secondary_color = "#00B4E0",
  logo_path = "path/to/logo.png",
  font_family = "Arial"
)
```

## Template Components

### CSS Customization

The template generates a custom CSS file with your branding:

```css
/* Custom styles for Quarto template */
:root {
  --primary-color: #0054AD;
  --secondary-color: #00B4E0;
}

.navbar {
  background-color: var(--primary-color);
}

.title {
  color: var(--primary-color);
}

a {
  color: var(--secondary-color);
}

body {
  font-family: Arial;
}
```

### YAML Configuration

The template includes a YAML configuration:

```yaml
format:
  html:
    theme: cosmo
    css: custom.css
    toc: true
    code-fold: true
    code-tools: true
    df-print: paged
    fig-width: 8
    fig-height: 6
    fig-format: png
    fig-dpi: 300
```

## Applying Templates

Apply your template to a Quarto document:

```{r eval=FALSE}
apply_template_to_report("reports/analysis.qmd", "company_template")
```

## Advanced Customization

### Custom Fonts

Use web fonts in your template:

```{r eval=FALSE}
create_quarto_template(
  template_name = "modern_template",
  font_family = "'Source Sans Pro', sans-serif",
  primary_color = "#2E5EAA"
)
```

Add the font link to your template:

```yaml
format:
  html:
    css: custom.css
    include-in-header:
      text: |
        <link href="https://fonts.googleapis.com/css2?family=Source+Sans+Pro:wght@400;600&display=swap" rel="stylesheet">
```

### Custom Layouts

Create custom layouts with CSS Grid:

```css
.custom-layout {
  display: grid;
  grid-template-columns: 1fr 2fr;
  gap: 2rem;
}

.sidebar {
  background: var(--primary-color);
  padding: 1rem;
  color: white;
}

.main-content {
  padding: 1rem;
}
```

### Interactive Elements

Add custom JavaScript for interactivity:

```js
// custom.js
document.addEventListener('DOMContentLoaded', function() {
  // Add custom behaviors
});
```

Include in your template:

```yaml
format:
  html:
    include-after-body:
      file: custom.js
```

## Best Practices

1. **Color Usage**
   - Use consistent color palette
   - Ensure sufficient contrast
   - Consider color blindness accessibility

2. **Typography**
   - Choose readable fonts
   - Maintain consistent hierarchy
   - Use appropriate font sizes

3. **Layout**
   - Keep layouts clean and consistent
   - Use white space effectively
   - Make navigation intuitive

4. **Responsive Design**
   - Test on different screen sizes
   - Use flexible layouts
   - Optimize images

## Template Organization

Recommended structure for template files:

```
reports/
└── templates/
    └── company_template/
        ├── _template.yml
        ├── custom.css
        ├── custom.js
        └── assets/
            └── logo.png
```

## Example Templates

### Executive Summary Template

```yaml
# _template.yml
format:
  html:
    theme: cosmo
    css: custom.css
    toc: false
    code-fold: true
    code-tools: false
    fig-width: 10
    fig-height: 6
    fig-format: png
    fig-dpi: 300
    include-in-header:
      file: header.html
```

### Technical Report Template

```yaml
# _template.yml
format:
  html:
    theme: cosmo
    css: custom.css
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    df-print: paged
    fig-width: 8
    fig-height: 6
    fig-format: png
    fig-dpi: 300
    include-in-header:
      file: header.html
```

## Integration with R Markdown

While Quarto is the preferred format, the templates can be adapted for R Markdown:

```yaml
output:
  html_document:
    theme: cosmo
    css: templates/company_template/custom.css
    toc: true
    code_folding: hide
```

## Next Steps

- Create templates for different report types
- Add custom JavaScript functionality
- Implement accessibility features
- Share templates across your organization 
