---
title: "Decision Tracking with thoth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Decision Tracking with thoth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE  # Don't evaluate code chunks by default
)
```


## Introduction

One of the challenges in bioinformatics and data analysis is documenting and tracking human decisions made during the analysis process. While code and data can be version controlled, the reasoning behind key analytical decisions often remains undocumented. The `thoth` package provides functionality to track these decisions in a structured way.

## Installation

To install the package from GitHub, first ensure you have all the required dependencies:

```{r installation}
# Install required dependencies if not already installed
required_packages <- c("usethis", "yaml", "cli", "digest", "rmarkdown", "tools")
for (pkg in required_packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}

# Install thoth from GitHub
if (!requireNamespace("devtools", quietly = TRUE)) {
  install.packages("devtools")
}
devtools::install_github("sebrauschert/thoth")
```

## Getting Started

First, let's load the package:

```{r setup}
library(thoth)
```

## Creating a Decision Tree

When starting a new analysis, initialize a decision tree to track your decisions:

```{r}
decision_file <- initialize_decision_tree(
  analysis_id = "RNA_seq_2024_01",
  analyst = "Jane Smith",
  description = "Differential expression analysis of treatment vs control samples"
)
```

## Recording Decisions

As you progress through your analysis, record key decisions and observations:

```{r}
# Record a quality control decision
record_decision(
  file_path = decision_file,
  check = "Sample-wise PCA clustering",
  observation = "Treatment samples cluster together except for sample T3",
  decision = "Exclude sample T3",
  reasoning = "T3 clusters with controls, likely sample swap",
  evidence = "plots/PCA_pre_filtering.pdf"
)

# Record a biological validation step
record_decision(
  file_path = decision_file,
  check = "Known pathway markers",
  observation = "Expected stress response genes upregulated",
  decision = "Results biologically plausible",
  reasoning = "Key marker genes show expected direction of change",
  evidence = "tables/marker_genes_expression.csv"
)
```

## Generating Methods Documentation

You can automatically generate a methods section from your decision tree:

```{r}
# Generate methods section
methods_text <- generate_methods_section(decision_file)
cat(methods_text)
```

## Exporting Decision Trees

Export your decision tree to various formats for sharing or documentation:

```{r}
# Export to markdown
export_decision_tree(decision_file, format = "md")

# Export to HTML (requires rmarkdown)
if (requireNamespace("rmarkdown", quietly = TRUE)) {
  export_decision_tree(decision_file, format = "html")
}
```

## Integration with Project Workflow

The decision tracking functionality integrates seamlessly with your analysis workflow:

1. Initialize a decision tree at the start of your analysis
2. Record decisions as you make them
3. Use the exported documentation for methods sections in papers
4. Keep the decision tree under version control

## Best Practices

Here are some recommended practices for using the decision tracking functionality:

1. **Be Consistent**: Record decisions as you make them, not after the fact
2. **Be Specific**: Include clear observations and reasoning
3. **Link Evidence**: Always link to supporting plots, tables, or other evidence
4. **Version Control**: Keep decision trees under version control
5. **Review**: Regularly review decisions with team members

## Conclusion

By tracking analytical decisions in a structured way, we can:

- Make the analysis process more transparent
- Facilitate knowledge transfer within teams
- Improve reproducibility
- Generate better documentation

The decision tracking functionality in `thoth` helps make human expertise as trackable as code, breaking free from the "90% automation trap" where crucial human judgment remains hidden.
