---
title: "Docker Setup with thoth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Docker Setup with thoth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`thoth` simplifies containerization by automatically creating a Docker environment that matches your local R setup. This ensures your analysis can be reproduced exactly as intended, regardless of where it runs.

## Automatic Setup

When creating a new project, `thoth` handles Docker configuration automatically:

```{r}
library(thoth)

create_analytics_project(
  "my_analysis",
  use_docker = TRUE
)
```

This creates two key files:

### 1. Dockerfile

```dockerfile
FROM rocker/rstudio:${R_VERSION}

# System dependencies
RUN apt-get update && apt-get install -y \
    python3-pip \
    && rm -rf /var/lib/apt/lists/*

# Install DVC
RUN pip3 install dvc

# Project setup
WORKDIR /project
COPY . /project/

# R package installation
RUN R -e 'install.packages("renv")'
RUN R -e 'renv::restore()'

# Set permissions
RUN chown -R rstudio:rstudio /project

CMD ["/init"]
```

### 2. docker-compose.yml

```yaml
services:
  rstudio:
    build: .
    ports:
      - "8787:8787"
    environment:
      - PASSWORD=rstudio
      - ROOT=TRUE
    volumes:
      - .:/project
    user: rstudio
```

## Using the Environment

### Starting RStudio Server

```bash
# Start the container
docker-compose up -d

# Access RStudio at http://localhost:8787
# Username: rstudio
# Password: rstudio

# Stop the container
docker-compose down
```

## Key Features

### 1. Version Matching

- Uses exactly the same R version as your local installation  
- Ensures perfect reproducibility across environments  
- Automatically handles system dependencies  

### 2. Development Integration

- Mounts your project directory  
- Real-time file synchronization  
- Preserves Git and DVC functionality  

### 3. Package Management

- Integrates with `renv` for dependency tracking  
- Automatically installs required packages  
- Maintains consistent package versions  

## Best Practices

### Project Organization

- Keep Docker files in version control  
- Document any customizations  
- Include Docker instructions in README  

### Development Workflow
```{r}
# 1. Create project with Docker
create_analytics_project("analysis", use_docker = TRUE)

# 2. Start container
system("docker-compose up -d")

# 3. Develop in RStudio Server
# Access at http://localhost:8787

# 4. Track dependencies automatically
# renv and Docker handle the rest
```

### Sharing
```bash
# Recipients only need to:
git clone <repository>
cd <repository>
docker-compose up -d
```

## Customization

You can customize the Docker setup by modifying:

1. **Dockerfile**: Add tools or system dependencies
```dockerfile
# Add custom system packages
RUN apt-get update && apt-get install -y \
    your-package-here

# Add R packages
RUN R -e 'install.packages("your-package")'
```

2. **docker-compose.yml**: Adjust container settings
```yaml
services:
  rstudio:
    ports:
      - "custom-port:8787"
    environment:
      - CUSTOM_VAR=value
```

## Next Steps

- Try the end-to-end example: `vignette("end-to-end-example")`  
- Learn about DVC integration: `vignette("dvc-tracking")`  
- Check [Docker documentation](https://docs.docker.com/) for advanced features  
