---
title: "Getting Started with thoth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with thoth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`thoth` is a comprehensive framework for setting up reproducible analytics projects in R. It integrates several best practices and tools:

- Data version control with DVC (through R functions)
- Containerization with Docker
- Dependency management with renv
- Customizable reporting with Quarto

This guide will help you get started with the basic features of `thoth`.

## Installation

You can install the development version of `thoth` from GitHub:

```{r}
# install.packages("devtools")
devtools::install_github("sebrauschert/thoth")
```

### Prerequisites

Before using `thoth`, ensure you have:

1. [DVC](https://dvc.org/doc/install) installed
2. [Docker](https://docs.docker.com/get-docker/) installed
3. [R](https://www.r-project.org/) (>= 4.1.0)
4. [Quarto](https://quarto.org/docs/get-started/) for report generation

## Creating a New Project

Create a new analytics project with all the necessary setup:

```{r}
library(thoth)
library(tidyverse)

# Create a new project with all features enabled
create_analytics_project(
  "my_project",
  use_dvc = TRUE,
  use_docker = TRUE,
  use_quarto = TRUE
)
```

This sets up a standardized project structure:

```
my_project/
├── .dvc/               # DVC configuration
├── .git/               # Git version control
├── .gitignore         # Git ignore patterns
├── .Rproj.user/      # RStudio project files
├── data/             
│   ├── raw/          # Raw data (tracked by DVC)
│   └── processed/    # Processed data (tracked by DVC)
├── R/                # R scripts
├── reports/          # Quarto reports
├── docker/           # Dockerfile and related files
├── renv/             # renv library and lockfile
└── README.md         # Project documentation
```

## Data Version Control with thoth

`thoth` provides seamless integration with DVC through tidyverse-style functions. Here's how to use it in your data science workflow:

### Reading and Processing Data

```{r}
# Example data processing pipeline
raw_data <- read_csv("data/raw/dataset.csv")

processed_data <- raw_data |>
  mutate(
    date = as.Date(date),
    value = as.numeric(value)
  ) |>
  filter(!is.na(value)) |>
  # Save processed data with DVC tracking
  write_csv_dvc(
    "data/processed/cleaned_data.csv",
    "Cleaned and processed raw data"
  )

# Continue with analysis
summary_stats <- processed_data |>
  group_by(category) |>
  summarise(
    mean_value = mean(value),
    sd_value = sd(value)
  ) |>
  # Save summary statistics with DVC tracking
  write_csv_dvc(
    "data/processed/summary_stats.csv",
    "Generated summary statistics"
  )
```

### Tracking Models and R Objects

```{r}
# Train a model
model <- lm(value ~ date + category, data = processed_data)

# Save model with DVC tracking
model |>
  write_rds_dvc(
    "models/linear_model.rds",
    "Trained linear regression model"
  )

# Make predictions and save results
predictions <- model |>
  predict(newdata = new_data) |>
  as_tibble() |>
  write_csv_dvc(
    "results/model_predictions.csv",
    "Generated model predictions"
  )
```

### Best Practices for Data Organization

```{r}
# Raw data pipeline
raw_survey_data |>
  write_csv_dvc(
    "data/raw/survey_2024.csv",
    "Added raw survey data for 2024"
  )

# Processed data pipeline
cleaned_survey_data |>
  write_csv_dvc(
    "data/processed/survey_2024_clean.csv",
    "Added cleaned survey data"
  ) |>
  create_features() |>
  write_csv_dvc(
    "data/processed/survey_2024_features.csv",
    "Added feature engineered dataset"
  )

# Model pipeline
final_model |>
  write_rds_dvc(
    "models/survey_2024_model.rds",
    "Final predictive model with accuracy: 0.92"
  )
```

## Docker Integration

`thoth` sets up a Docker environment with R, RStudio Server, and all necessary tools:

```bash
# Build and start the container
docker compose up -d

# Access RStudio Server at http://localhost:8787
# Username: rstudio
# Password: rstudio
```

The Docker setup includes:
- R and RStudio Server
- DVC for data version control
- All system dependencies
- Automatic package installation via renv

## Dependency Management

`thoth` uses `renv` for dependency management:

```{r}
# Install a new package
renv::install("tidymodels")

# Update lockfile
renv::snapshot()

# Restore project environment
renv::restore()
```

## Quarto Reports

Create beautiful reports with custom templates:

```{r}
# Create a new template
create_quarto_template(
  "company_template",
  primary_color = "#0054AD",
  secondary_color = "#00B4E0",
  logo_path = "path/to/logo.png"
)

# Apply template to a report
apply_template_to_report(
  "reports/analysis.qmd",
  "company_template"
)
```

## Best Practices

1. **Data Organization**
   ```{r}
   # Keep raw data separate
   raw_data |>
     write_csv_dvc(
       "data/raw/experiment_data.csv",
       "Raw experimental data"
     )
   
   # Store processed data with clear naming
   processed_data |>
     write_csv_dvc(
       "data/processed/experiment_cleaned.csv",
       "Cleaned experimental data"
     )
   
   # Save models with version info
   model |>
     write_rds_dvc(
       "models/xgboost_v2.rds",
       "XGBoost model v2 (accuracy: 0.92)"
     )
   ```

2. **Error Handling**
   ```{r}
   # Safe writing function
   safe_write <- function(data, path, message) {
     tryCatch(
       write_csv_dvc(data, path, message),
       error = function(e) {
         warning("Failed to save data: ", e$message)
         NULL
       }
     )
   }
   
   # Using purrr::safely
   safe_write_dvc <- safely(write_csv_dvc)
   result <- safe_write_dvc(data, "data/output.csv", "Saved output")
   ```

3. **Documentation**
   - Document analysis steps in reports
   - Keep README.md up to date
   - Add comments to complex code
   - Use descriptive commit messages

4. **Reproducibility**
   - Use Docker for environment consistency
   - Track dependencies with renv
   - Version control everything
   - Write automated tests

## Next Steps

- Read the [DVC Integration Guide](dvc-integration.html) for advanced usage
- Learn about [Docker Setup](docker-setup.html) for customizing your environment
- Explore [Custom Templates](custom-templates.html) for report styling
- Check out the [function reference](../reference/index.html) for detailed documentation

## Getting Help

If you encounter issues or have questions:

1. Check the [package documentation](https://sebrauschert.github.io/thoth/)
2. File issues on [GitHub](https://github.com/sebrauschert/thoth/issues)
3. Read the function documentation with `?function_name` 