---
title: "Git Integration with thoth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Git Integration with thoth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

`thoth` provides a tidyverse-friendly interface to Git operations, allowing you to manage version control directly from your R environment. This integration streamlines the version control workflow while maintaining the power and flexibility of Git.

## Core Git Functions

### Repository Status and Changes

Monitor and manage changes in your repository:

```{r}
library(thoth)

# Check repository status
git_status()

# Stage changes
git_add("analysis.R")                                    # Single file
git_add(c("data/processed/results.csv", "README.md"))   # Multiple files
git_add(".", force = FALSE)                             # All changes

# Commit changes
git_commit("Add initial analysis")                      # Commit staged changes
git_commit("Update results", all = TRUE)                # Stage and commit all changes
```

### Branch Management

Create and manage branches efficiently:

```{r}
# View branches
git_branch_list()                    # Local branches
git_branch_list(all = TRUE)          # Include remote branches

# Create and switch branches
git_branch("feature/analysis")       # Create new branch
git_checkout("main")                 # Switch branch
git_checkout("feature/fix",          # Create and switch
             create = TRUE)
```

### Remote Operations

Synchronize with remote repositories:

```{r}
# Sync changes
git_pull()                          # Get latest changes
git_push()                          # Push local changes

# Work with specific remotes/branches
git_pull("origin", "main")
git_push("origin", "feature/analysis")
```

## Integration with DVC

`thoth`'s Git functions work seamlessly with DVC for complete version control:

```{r}
# Track data with DVC
dvc_add("data/raw/dataset.csv")

# Version control the tracking file
git_add("data/raw/dataset.csv.dvc")
git_commit("Add dataset tracking")

# Sync everything
git_push()      # Push Git changes
dvc_push()      # Push data to remote storage
```

## Best Practices

### 1. Commit Structure  

- Make focused, single-purpose commits  
- Write clear, present-tense commit messages  
- Include context when necessary  

### 2. Branch Management  

- Use feature branches for new work  
- Keep the main branch stable  
- Clean up merged branches  

### 3. Workflow Integration
```{r}
# Example analysis workflow
git_checkout("feature/new-analysis", create = TRUE)

# Do analysis work...

git_add("R/analysis.R")
git_commit("Add data processing function")
git_push()
```

## Common Workflows

### Feature Development

```{r}
# Start new feature
git_checkout("main")
git_pull()
git_checkout("feature/analysis", create = TRUE)

# Work on feature...
git_add(".")
git_commit("Implement new analysis")
git_push()
```

### Quick Fixes

```{r}
# Create fix branch
git_checkout("main")
git_checkout("fix/issue-123", create = TRUE)

# Make changes...
git_add("R/functions.R")
git_commit("Fix calculation error")
git_push()
```

## Next Steps  

- Explore the end-to-end example: `vignette("end-to-end-example")`  
- Learn about DVC integration: `vignette("dvc-tracking")`  
- Check the [Git documentation](https://git-scm.com/doc) for advanced features  
