---
title: "Docker Setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Docker Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction to Docker with R

Docker provides a way to package your R environment, dependencies, and code into a reproducible container. `toth` automatically sets up a Docker configuration that includes R, RStudio Server, and all necessary tools for reproducible analytics.

## Understanding the Default Setup

When you create a new project with `toth`, it generates two key Docker files:

### Dockerfile

```dockerfile
FROM rocker/verse:latest

# Install system dependencies
RUN apt-get update && apt-get install -y \
    python3-pip \
    && rm -rf /var/lib/apt/lists/*

# Install DVC
RUN pip3 install dvc

# Create working directory
WORKDIR /project

# Copy project files
COPY . /project/

# Install R packages
RUN R -e 'renv::restore()'

# Command to keep the container running
CMD ["/bin/bash"]
```

### docker-compose.yml

```yaml
services:
  rstudio:
    build: .
    ports:
      - "8787:8787"
    environment:
      - PASSWORD=rstudio
    volumes:
      - .:/project
```

## Basic Usage

### Starting the Environment

```bash
# Build and start the container
docker compose up -d

# Access RStudio Server at http://localhost:8787
# Username: rstudio
# Password: rstudio
```

### Stopping the Environment

```bash
docker compose down
```

## Customizing the Environment

### Adding System Dependencies

Modify the Dockerfile to include additional system packages:

```dockerfile
RUN apt-get update && apt-get install -y \
    python3-pip \
    libgdal-dev \
    libproj-dev \
    && rm -rf /var/lib/apt/lists/*
```

### Adding R Packages

1. **Using renv** (Recommended):
   ```r
   # In R
   renv::install("package_name")
   renv::snapshot()
   ```

2. **Direct Installation** (Dockerfile):
   ```dockerfile
   RUN R -e 'install.packages("package_name")'
   ```

### Customizing RStudio Settings

Add custom RStudio preferences:

```dockerfile
COPY rstudio-prefs.json /home/rstudio/.config/rstudio/rstudio-prefs.json
```

## Advanced Configuration

### Environment Variables

Use a `.env` file for configuration:

```bash
# .env
POSTGRES_PASSWORD=secret
AWS_ACCESS_KEY_ID=your_key
AWS_SECRET_ACCESS_KEY=your_secret
```

Update docker-compose.yml:

```yaml
services:
  rstudio:
    env_file:
      - .env
```

### Adding Services

Extend docker-compose.yml to include additional services:

```yaml
services:
  rstudio:
    # ... existing config ...
  
  postgres:
    image: postgres:latest
    environment:
      - POSTGRES_PASSWORD=${POSTGRES_PASSWORD}
    volumes:
      - postgres_data:/var/lib/postgresql/data

volumes:
  postgres_data:
```

## Best Practices

1. **Security**
   - Never commit sensitive information in Dockerfile
   - Use .env files for secrets (and add to .gitignore)
   - Regularly update base images

2. **Performance**
   - Use .dockerignore to exclude unnecessary files
   - Layer your Dockerfile efficiently
   - Consider multi-stage builds for complex setups

3. **Development Workflow**
   - Use volumes for development
   - Set up proper user permissions
   - Configure Git to ignore Docker-specific files

## Troubleshooting

Common issues and solutions:

1. **Permission Issues**
   ```bash
   # Fix file ownership
   sudo chown -R 1000:1000 .
   ```

2. **Port Conflicts**
   ```yaml
   # Change port mapping in docker-compose.yml
   ports:
     - "8788:8787"  # Use different host port
   ```

3. **Memory Issues**
   ```yaml
   services:
     rstudio:
       mem_limit: 4g
       mem_reservation: 2g
   ```

## Integration with toth

`toth` provides convenient setup and configuration:

```{r eval=FALSE}
library(toth)

# Create project with Docker support
create_analytics_project(
  "my_project",
  use_docker = TRUE
)

# This sets up:
# - Dockerfile with R and system dependencies
# - docker-compose.yml for easy orchestration
# - .dockerignore for proper file exclusion
# - Default RStudio configuration
```

## Next Steps

- Explore Docker's [official documentation](https://docs.docker.com)
- Learn about Docker Compose features
- Set up continuous integration with Docker
- Create custom base images for your needs 