---
title: "DVC with toth"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DVC with toth}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Introduction

The `toth` package provides seamless integration with DVC (Data Version Control) for tracking data files in your R analytics projects. This guide demonstrates how to use DVC through R functions, making it easy to version control your data alongside your code.

## Project Setup

The `toth` package automatically configures DVC when creating a new project:

```{r setup}
library(toth)
library(tidyverse)

# Create a new project with DVC enabled
create_analytics_project(
  "my_project",
  use_dvc = TRUE,
  git_init = TRUE
)
```

This sets up:
- DVC initialization
- Appropriate `.gitignore` rules
- Standard project structure
- DVC configuration files

## Basic Usage

### Simple Data Tracking

The most basic way to track files is to use `write_csv_dvc()` or `write_rds_dvc()`:

```{r}
# Create example data
data <- tibble(
  x = 1:5,
  y = letters[1:5]
)

# Write and track CSV file
data |>
  write_csv_dvc(
    "data/processed/example.csv",
    "Added processed data"
  )

# Track R objects
model <- lm(mpg ~ wt, data = mtcars)
model |>
  write_rds_dvc(
    "models/linear_model.rds",
    "Saved linear model"
  )
```

### Data Processing Pipelines

Integrate DVC tracking into your data processing workflows:

```{r}
# Example data pipeline
raw_data |>
  # Clean data
  mutate(
    date = as.Date(date),
    value = as.numeric(value)
  ) |>
  # Save intermediate results
  write_csv_dvc(
    "data/interim/cleaned.csv",
    "Saved cleaned data"
  ) |>
  # Continue processing
  group_by(category) |>
  summarise(
    mean_value = mean(value),
    sd_value = sd(value)
  ) |>
  # Save final results
  write_csv_dvc(
    "data/processed/summary.csv",
    "Saved summary statistics"
  )
```

### Tracking Multiple Files

You can track multiple files in sequence:

```{r}
# Split data by category
data |>
  group_split(category) |>
  walk2(
    unique(data$category),
    ~write_csv_dvc(
      .x,
      file.path("data/processed", paste0(.y, ".csv")),
      paste("Saved data for category", .y)
    )
  )
```

## Best Practices

### File Organization

```{r}
# Raw data
raw_data |>
  write_csv_dvc(
    "data/raw/survey_2024.csv",
    "Added raw survey data"
  )

# Processed data
processed_data |>
  write_csv_dvc(
    "data/processed/survey_2024_clean.csv",
    "Added cleaned survey data"
  )

# Analysis results
results |>
  write_csv_dvc(
    "results/survey_2024_analysis.csv",
    "Added analysis results"
  )

# Models with version info
model |>
  write_rds_dvc(
    "models/xgboost_v2.rds",
    "Updated XGBoost model with hyperparameter tuning (accuracy: 0.92)"
  )
```

### Error Handling

```{r}
# Using tryCatch for error handling
safe_write <- function(data, path, message) {
  tryCatch(
    write_csv_dvc(data, path, message),
    error = function(e) {
      warning("Failed to save data: ", e$message)
      NULL
    }
  )
}

# Using purrr::safely
safe_write_dvc <- safely(write_csv_dvc)
result <- safe_write_dvc(data, "data/output.csv", "Saved output")

# Check system requirements
check_system_requirements()
```

## Advanced Usage

### Conditional Tracking

Track files based on conditions:

```{r}
process_and_track <- function(data, threshold = 0.9) {
  results <- analyze_data(data)
  
  if (results$accuracy > threshold) {
    results |>
      write_csv_dvc(
        "results/high_accuracy.csv",
        sprintf("Results with accuracy %.2f", results$accuracy)
      )
  } else {
    results |>
      write_csv_dvc(
        "results/low_accuracy.csv",
        sprintf("Results with accuracy %.2f", results$accuracy)
      )
  }
}
```

### Tracking with Metadata

Include relevant metadata in commit messages:

```{r}
save_model_version <- function(model, version, metrics) {
  model |>
    write_rds_dvc(
      sprintf("models/model_v%s.rds", version),
      sprintf(
        "Model v%s: accuracy=%.3f, f1=%.3f",
        version, metrics$accuracy, metrics$f1
      )
    )
}
```

## Integration with Other Tools

### renv Integration

```{r}
# Save environment with model
list(
  model = model,
  packages = renv::dependencies()
) |>
  write_rds_dvc(
    "models/model_with_deps.rds",
    "Saved model with package dependencies"
  )
```

### Collaboration Workflow

```{r}
# After pulling new code with git pull:
new_data |>
  write_csv_dvc(
    "data/raw/new_experiment.csv",
    "Add new experiment data"
  ) |>
  process_data() |>
  write_csv_dvc(
    "data/processed/new_experiment_processed.csv",
    "Add processed experiment data"
  ) |>
  train_model() |>
  write_rds_dvc(
    "models/new_experiment_model.rds",
    "Add model for new experiment"
  )
```

## Next Steps

- Check out the [Getting Started guide](getting-started.html) for basic package usage
- Learn about [Docker Setup](docker-setup.html) for containerized environments
- Explore [Custom Templates](custom-templates.html) for report styling
- Visit the [function reference](../reference/index.html) for detailed documentation

## Getting Help

- Read the function documentation with `?dvc_track`, `?write_csv_dvc`, or `?write_rds_dvc`
- File issues on [GitHub](https://github.com/sebrauschert/toth/issues)
- Check the [package website](https://sebrauschert.github.io/toth) for updates 